---
title: "Comparing Health Insurance Exchange Data, 2014-2018"
author: "Nate Apathy"
date: "2/13/2018"
output: html_document
---
```{r packages, include=FALSE}
library(foreign) # this allows for pulling in the data from the website
library(survey) # this is for analyzing survey data
library(dplyr)
library(expss)
library(car)
library(ggplot2)
library(stargazer)
library(MASS)
library(pscl)
library(tidyverse)
library(geofacet)
library(reshape2)
library(ggthemes)
library(psych)
library(skimr)
library(doBy)
```

# What is HIX Compare?
[HIX compare](https://www.hixcompare.org/) is the website where files for the ACA health insurance exchange marketplace plans are stored and available for download. This allows researchers to look at insurance plans across both the individual and group markets longitudinally.

# What are we going to do?
We're going to use these data sets to learn some lessons about how to bring data into R, clean it up, format it, merge it, subset it, and some other stuff. We will work only with the [individual market](https://www.hixcompare.org/individual-markets.html) plan files from 2014 through 2018. If you care about the small group market plans, those are available, too.

# Getting the data
First, we have to learn about what is in the files, which can be done with the [documentation file](https://data.hixcompare.org/hix_compare_documentation_2018_02_02.pdf).

## Pulling the data into R
Websites for the files we are using:
- [2014 zip file](https://data.hixcompare.org/previous_versions/individual_2014_2018_02_02.zip)
- [2015 zip file](https://data.hixcompare.org/previous_versions/individual_2015_2018_02_02.zip)
- [2016 zip file](https://data.hixcompare.org/previous_versions/individual_2016_2018_02_02.zip)
- [2017 zip file](https://data.hixcompare.org/previous_versions/individual_2017_2018_02_02.zip)
- [2018 zip file](https://data.hixcompare.org/previous_versions/individual_2018_2018_02_02.zip)

## Downloading and saving the data as an .Rdata file
### DO NOT RUN THIS IF YOU ALREADY HAVE THE .Rdata FILES
- You only have to do this part once, to download the raw data and then save it as an .Rdata file.
- Then, you keep the .Rdata file in your project directory and do a "load" function to retrieve the data in the future (second code chunk).
```{r downloading, eval=FALSE}
# there's a way to loop this and write it as a function
# but this is easier to understand what's going on and see the process

# 2014
download.file("https://data.hixcompare.org/previous_versions/individual_2014_2018_02_02.zip", 
              temp <- tempfile())
unzipped_file <- unzip(temp)
hix14 <- read_csv(unzipped_file[1])
unlink(temp)  # Unlink to delete temporary file

# 2015
download.file("https://data.hixcompare.org/previous_versions/individual_2015_2018_02_02.zip", 
              temp <- tempfile())
unzipped_file <- unzip(temp)
hix15 <- read_csv(unzipped_file[1])
unlink(temp)  # Unlink to delete temporary file

# 2016
download.file("https://data.hixcompare.org/previous_versions/individual_2016_2018_02_02.zip", 
              temp <- tempfile())
unzipped_file <- unzip(temp)
hix16 <- read_csv(unzipped_file[1])
unlink(temp)  # Unlink to delete temporary file

# 2017
download.file("https://data.hixcompare.org/previous_versions/individual_2017_2018_02_02.zip", 
              temp <- tempfile())
unzipped_file <- unzip(temp)
hix17 <- read_csv(unzipped_file[1])
unlink(temp)  # Unlink to delete temporary file

# 2018
download.file("https://data.hixcompare.org/previous_versions/individual_2018_2018_02_02.zip", 
              temp <- tempfile())
unzipped_file <- unzip(temp)
hix18 <- read_csv(unzipped_file[1])
unlink(temp)  # Unlink to delete temporary file

# clean up your environment by removing these things we don't really need
rm("temp","theurl","unzipped_file")

# Save all as .Rdata files for faster loading later
# KEEP THESE AS YOUR RAW DATA FILES - NEVER EDIT THESE
save(hix14, file="hix14.Rdata")
save(hix15, file="hix15.Rdata")
save(hix16, file="hix16.Rdata")
save(hix17, file="hix17.Rdata")
save(hix18, file="hix18.Rdata")
```

# Loading the Data Sets
```{r loading}
# In future loading, just use these two commands
# the files should already be in our project directory
load(file="hix14.Rdata")
load(file="hix15.Rdata")
load(file="hix16.Rdata")
load(file="hix17.Rdata")
load(file="hix18.Rdata")
```

# Merging the data

Now we have both of the datasets saved, and the first step is to merge them. But before we do that we need to check to make sure the columns all match up between the data sets. There are 503 variables in each file, but they might be in different orders, or contain different data, or be named different things. So let's check.

```{r checkcols}
# turn the column names and information about them into an object
# use the "structure" function
hix14cols <- colnames(hix14)
hix15cols <- colnames(hix15)
hix16cols <- colnames(hix16)
hix17cols <- colnames(hix17)
hix18cols <- colnames(hix18)

# then compare them all to each other
# need to see exactly what differs between them
# takes two objects and checks to see if they are identical
identical(hix14cols,hix15cols)
identical(hix15cols,hix16cols)
identical(hix16cols,hix17cols)
identical(hix17cols,hix18cols)
```

So all the columns match; great. We just merge them all vertically, right? Just stack one on top of the other?

Are you sure?

Yes, we can do this because they already all have a "YEAR" variable to tell them apart. We're also super lucky here in that they are all the same amount of columns in the same order. So we can do a very simple "row bind" which stacks them on top of each other.

```{r merge}
# i'm sure there are better ways to do this with the tidyverse
# i just don't know how yet ;-)
hix1415 <- rbind(hix14,hix15)
hix1617 <- rbind(hix16,hix17)
hix1417 <- rbind(hix1415,hix1617)
hix1418 <- rbind(hix1417,hix18)

# clean up the ones you don't need anymore
rm("hix1415","hix1617","hix1417")

# drop child-only and CSRs
hix1418 <- hix1418[hix1418$CHILDONLY==0 & hix1418$CSR==0,]
```

# Defining Unique Plans
HIX Compare has a suggested methodology for defining "unique" plans to which we should pay heed.

Also, this is a good lesson in data management - identifying your unit of analysis and making sure your data reflects that.

```{r uniqueplans, results="hide"}
# we need to find the records for which the following fields are duplicated
# year st carrier metal plantype planmarket networkid
# we are going to add area to this set, because we want to analyze at the area level (col #6)
# ab_copayinn_tiers *thru* rh_coinsoutofneta
# sp_copayinn_tiers *thru* tehboutofnetfamilymoopa
# this leaves out skilled nursing columns (there are 14 of them) because they are an EHB that can be changed by a rider
# we're going to use the UNIQUE field once we compress the data to hold our unique ID for each plan that remains
# check for duplications T/F
duplicated(hix1418[,c(2,5:7,9,10,17,19,20:383,398:503)])
# returns a logical vector that matches with the rows that are duplicated by the hixcompare rules

colnames(hix1418[1:3,c(2,5:7,9,10,17,19,20:383,398:503)])
# yep, those fields are correct

# now we can just use the duplicated() function and its arguments as our subset for rows

un_hix1418 <- hix1418[!duplicated(hix1418[,c(2,5:7,9,10,17,19,20:383,398:503)]),]
# this cuts us down to 139,516 unique plans, down from 168,177

# now we can generate our unique identifier in the UNIQUE field
un_hix1418$UNIQUE <- 1:length(un_hix1418$UNIQUE)
```

# Understanding the Data
## READ THE DATA DICTIONARY
Find out what is in the data dictionary so that you understand it before we go exploring. The data dictionary is in the github repository for reference and can be found on the hixcompare.org website for the individual market files.
- need to know if your project requires individual or small group plan data


```{r datadict, results="hide"}
colnames(un_hix1418)
# Common fields we will always want to keep: 1-19 and 440-503
# ER: cols 90-103 (14 columns)
```

Now that we have the datasets completely merged... let's take a look at this data.

```{r dimensions}
# what are the dimensions of this data frame?
dim(un_hix1418)
# 139,516 observations over 503 variables

# how many cells/potential data points is that?
dim(un_hix1418)[1]*dim(un_hix1418)[2]
# 70,176,548
```

So we have ~70 million observations over 5 years of insurance plan data.

This will probably need some cleanup. 

# Exploring the data
```{r}
# table(un_hix1418$CARRIER,un_hix1418$YEAR)
# this table is huge... so let's not worry about it.

table(un_hix1418$YEAR)

table(un_hix1418$YEAR,un_hix1418$METAL)
```

# Subsetting
## For Analysis
- within indiana, for things that we are bad at (pre-natal care), is there meaningful variation within tiers
    - looking at variation within plans & services; is there variation w/in areas and tiers?
    - let's start by looking just at 2018 in indiana, within silver plans

```{r subset_analysis, warning=FALSE}
in18s <- un_hix1418[un_hix1418$ST=="IN" & un_hix1418$METAL=="Silver" & un_hix1418$YEAR=="2018",]
# only 270 silver plans in indiana during this time
# but that's still a lot!
# let's see how many per area
table(in18s$ST,in18s$AREA)

# and let's now just look at the plans in IN10 (20 of them) and their ER cost sharing
in18s10er <- in18s[in18s$AREA=="IN10",c(1:19, # pricing and metadata
                                        90:103, # ER-specific fields
                                        440:503)] # value fields
# area 10 and pre/post natal care fields
in18s10pn <- in18s[in18s$AREA=="IN10",c(1:19, # pricing and metadata
                                        328:341, # pre/postnatal specific fields
                                        440:503)] # value fields

# all of indiana pre/post natal
in18spn <- in18s[,c(1:19, # pricing and metadata
                    328:341, # pre/postnatal specific fields
                    440:503)] # value fields

skim(in18spn)

ggplot(in18spn,aes(x=in18spn$TEHBInnTier1FamilyMOOPA,y=in18spn$TEHBDedInnTier1FamilyA)) +
  geom_point(alpha=0.3,aes(color=in18spn$CARRIER, size=in18spn$PREMC2C30))

ggplot(in18s,aes(x=PREMC2C30,y=TEHBDedInnTier1FamilyA)) +
  geom_point(alpha=0.3,aes(color=CARRIER, size=PREMC2C30)) +
  labs(x="Premium for 2 individuals (age 30) and 2 children (aged 0-14)",
       y="Integrated In-Network Deductible - Family")

in18 <- un_hix1418[un_hix1418$ST=="IN" & un_hix1418$YEAR=="2018",]

skim(in18)

unique(in18$CARRIER)

ggplot(in18,aes(x=PREMC2C30,y=TEHBInnTier1FamilyMOOPA)) +
  geom_point(alpha=0.3,aes(color=METAL,size=TEHBDedInnTier1FamilyA)) +
  labs(x="Premium for 2 individuals (age 30) and 2 children (aged 0-14)",
       y="Integrated In-Network Max OOP - Family") +
  facet_wrap(~AREA)
```


## For Visualization
Let's look at trends in the average silver price for 27 year olds and 50 year olds by state, then by area.

```{r silvervis, warning=FALSE, comment=FALSE}
# subset to silvers only
silverdf <- un_hix1418[un_hix1418$METAL=="Silver", #rows
                    c("YEAR","ST","PLANID","AREA","PREMI27","PREMI50","CSR")] # columns
# beauty of subsetting is that you can just overwrite your data frame object with any additional columns if you realize later on you need more data
# the vast majority of plans are silver plans because that's how benchmarking is done

avgprems <- aggregate(silverdf,list(silverdf$YEAR,silverdf$ST),mean)[,c(1,2,7,8)]
colnames(avgprems) <- c("yr","state","prem27","prem50")
# we have to convert it to a format that ggplot likes
avgprems <- melt(avgprems,id.vars=c("yr","state"))
avgprems$variable <- ifelse(avgprems$variable=="prem27","avg 27 year old", "avg 50 year old")

ggplot(avgprems[avgprems$state=="AK",],aes(x=yr,group=variable,color=variable)) + geom_line(aes(y=value))
# what i want is this line graph for every state comparing premiums for a 27 year old over time with that of a 50 year old over time.

ggplot(avgprems, aes(variable, x=yr, y=value)) +
  geom_line(aes(group=variable, color=variable)) +
  theme_bw() +
  facet_geo(~ state, grid = "us_state_grid2") +
  scale_x_continuous(labels = function(x) paste0("'", substr(x, 3, 4))) +
  labs(title = "Average Silver Premiums for 27-year-olds & 50-year-olds, 2014-2018",
    caption = "Data Source: hixcompare.org",
    x = "Year",
    y = "Average Silver Plan Monthly Premium",
    color="Age Group") +
  theme_fivethirtyeight()
```


# Plotting
For example, if we're interested in looking at deductibles across plan metal tier by year, we will need to make sure those three variables are formatted correctly.

- Metal tier = "METAL"
- Premium for 27-year-old = "PREMI27"" (will filter out child only plans)
    - CHILDONLY = 0
- Year = "YEAR"

```{r formatfields}
plottest <- un_hix1418[un_hix1418$CHILDONLY==0,c("YEAR","CARRIER","PLANID","PREMI27","METAL")]
```

# Some quick plots
```{r plots1}
ggplot(data=plottest, mapping = aes(x=plottest$YEAR)) + 
  geom_bar(aes(fill=plottest$METAL))
```






