---
title: "hsR Session 4"
author: "Nate Apathy"
date: "2.23.2018"
output:
  html_document:
    df_print: paged
subtitle: Data Management II & Visualization for Exploratory Data Analysis
---
```{r packages, include=FALSE}
# install.packages()
library(VIM)
library(skimr)
library(tidyverse)
library(knitr)
library(RColorBrewer)
library(ggthemes)
```

# Review of Last Week's Activities

Each of you had a section to do something with the un_hix1418 dataset we created

- Some things that may come in handy:

    - A few of these will require the data dictionary as a reference
    - Keep in mind that subsetting by column number is much easier when fields have unweildy names and there are lots of them
    - You'll only need the first 19 columns for all the steps below
    - colnames() can help identify the number of the column you are trying to find
    - the syntax for subsetting is dataframename[rows,columns]
    - table() can help with cross-tabs/counts of variable pairs
    - length() counts how many elements are in a given object
    - skim(dataframename) can be very useful. try it! no need to create an object, just look at it in the Console output
    - you can subset within other functions without changing the object you are subsetting (if you don't overwrite it)
    - example: length(dataframename[dataframename$column1==1 & dataframename$column2==4,])
    - will count the number of observations that match your criteria (like filtering in excel)

## Subsetting & Exploring Data

```{r loaddfs}
load(file="un_hix1418.Rdata")
```

1. Subset the data frame down to the 2016 plans from two states of your choice (create a new data frame object)

```{r one}
alnm16 <- un_hix1418[(un_hix1418$ST=="AL" | un_hix1418$ST=="NM") & un_hix1418$YEAR=="2016",]

alnm16d <- un_hix1418 %>% filter(ST %in% c("AL","NM") & YEAR==2016)
```

2. How many plans are in each of the metallic tiers for each state?

```{r}
table(alnm16$METAL,alnm16$ST)
```

3. Which area (from the state first in alphabetical order) had the most silver plan options in 2016?

```{r}
table(alnm16[alnm16$ST=="AL" & alnm16$METAL=="Silver",]$AREA,alnm16[alnm16$ST=="AL" & alnm16$METAL=="Silver",]$ST)
```

4. Which state had the higher average premium for a 50-year-old purchasing a silver plan? What was that amount?

```{r}
mean(alnm16[alnm16$ST=="AL" & alnm16$METAL=="Silver",]$PREMI50)
mean(alnm16[alnm16$ST=="NM" & alnm16$METAL=="Silver",]$PREMI50)
```

5. What was the most common plan type in area 3 of the second state?

```{r}
table(alnm16[alnm16$ST=="NM",]$PLANTYPE,alnm16[alnm16$ST=="NM",]$AREA)
# type 2, which is HMO (from data dictionary)
```

6. Among HMO plans, what is the average premium for a 27-year-old?

```{r}
mean(alnm16[alnm16$PLANTYPE=="2",]$PREMI27)
```



# Visualizing Data Missingness
```{r}
table(un_hix1418$ST,un_hix1418$YEAR)
hixsamp <- filter(un_hix1418, YEAR %in% c("2016","2017") & ST %in% c("CA","MA")) %>%
  select(1:19,90:103)
colnames(hixsamp)[20:33]
aggr(hixsamp, prop = T, numbers=TRUE)
skim(hixsamp)
unique(hixsamp$CARRIER)
table(hixsamp$ER_CoinsInnTier1A,useNA = "always")
```

# Visualization of Data Summaries with the Tidyverse
```{r}
avg_prem_example <- un_hix1418 %>% group_by(YEAR, ST, METAL) %>%
  summarize(avg_27_prem = mean(PREMI27),
            avg_50_prem = mean(PREMI50),
            avg_fam4_prem = mean(PREMC2C30)) %>%
  mutate(avg_prem = (avg_27_prem + avg_50_prem + avg_fam4_prem)/3,
         METAL = factor(METAL,levels = c("Catastrophic","Bronze","Silver","Gold","Platinum")))

avg_prem_example


ggplot(avg_prem_example, aes(x=as.factor(YEAR))) + geom_bar(aes(fill=METAL))

ggplot(avg_prem_example, aes(x=as.factor(YEAR),y=avg_prem)) + geom_boxplot(aes(fill=METAL)) +
  labs(x="Year", y="Average Premium (state avg), 27-year old",
       title="Average Premiums by Metal Tier Over Time")

ggplot(avg_prem_example, aes(x=as.factor(YEAR))) + 
  geom_boxplot(aes(alpha=0.2,y=avg_27_prem)) +
  facet_wrap(~METAL)
```

# Class Exercise: Summarizing Data by State and Year

Question: On average, how many plans of each metallic tier were offered in each area of each state over the years?

```{r plancounts, include=FALSE}
plancounts <- un_hix1418 %>% group_by(AREA,ST,YEAR,METAL) %>% 
  summarize(plansinstate = n(), avg27prem = mean(PREMI27)) %>%
  group_by(YEAR,ST,METAL) %>%
  summarize(areaplans = mean(plansinstate), avg27prem = mean(avg27prem))

ggplot(plancounts, aes(x=YEAR,y=areaplans)) + geom_line(aes(color=METAL)) + facet_wrap(~ST)

ggplot(plancounts,aes(x=areaplans,y=avg27prem)) + geom_point(aes(color=METAL))
```

```{r saveplancounts}
save(plancounts, file="plancounts.Rdata")

```

