---
title: "hsR Session 4"
author: "Nate Apathy"
date: "2.23.2018"
output:
  html_document:
    df_print: paged
subtitle: Data Management II & Visualization for Exploratory Data Analysis
---
```{r packages, include=FALSE}
# install.packages()
library(VIM)
library(skimr)
library(tidyverse)
```

# Review of Last Week's Activities

Each of you had a section to do something with the un_hix1418 dataset we created

- Some things that may come in handy:

    - A few of these will require the data dictionary as a reference
    - Keep in mind that subsetting by column number is much easier when fields have unweildy names and there are lots of them
    - You'll only need the first 19 columns for all the steps below
    - colnames() can help identify the number of the column you are trying to find
    - the syntax for subsetting is dataframename[rows,columns]
    - table() can help with cross-tabs/counts of variable pairs
    - length() counts how many elements are in a given object
    - skim(dataframename) can be very useful. try it! no need to create an object, just look at it in the Console output
    - you can subset within other functions without changing the object you are subsetting (if you don't overwrite it)
    - example: length(dataframename[dataframename$column1==1 & dataframename$column2==4,])
    - will count the number of observations that match your criteria (like filtering in excel)

## Subsetting & Exploring Data

```{r loaddfs}
load(file="un_hix1418.Rdata")
```

1. Subset the data frame down to the 2016 plans from two states of your choice (create a new data frame object)

```{r one}
alnm16 <- un_hix1418[(un_hix1418$ST=="AL" | un_hix1418$ST=="NM") & un_hix1418$YEAR=="2016",]
```

2. How many plans are in each of the metallic tiers for each state?

```{r}
table(alnm16$METAL,alnm16$ST)
```

3. Which area (from the state first in alphabetical order) had the most silver plan options in 2016?

```{r}
table(alnm16[alnm16$ST=="AL" & alnm16$METAL=="Silver",]$AREA,alnm16[alnm16$ST=="AL" & alnm16$METAL=="Silver",]$ST)
```

4. Which state had the higher average premium for a 50-year-old purchasing a silver plan? What was that amount?

```{r}
mean(alnm16[alnm16$ST=="AL" & alnm16$METAL=="Silver",]$PREMI50)
mean(alnm16[alnm16$ST=="NM" & alnm16$METAL=="Silver",]$PREMI50)
```

5. What was the most common plan type in area 3 of the second state?

```{r}
table(alnm16[alnm16$ST=="NM",]$PLANTYPE,alnm16[alnm16$ST=="NM",]$AREA)
# type 2, which is HMO (from data dictionary)
```

6. Among HMO plans, what is the average premium for a 27-year-old?

```{r}
mean(alnm16[alnm16$PLANTYPE=="2",]$PREMI27)
```

# Visualizing Data Missingness
```{r}
hixsamp <- un_hix1418[1:500,1:19]
aggr(hixsamp, prop = F, numbers=TRUE)
skim(hixsamp)
```

