---
title: "hsR Session 5"
author: "Nate Apathy"
date: "3.2.2018"
output:
  html_document:
    df_print: paged
subtitle: Introduction to Basic Statistics in R
---
```{r packages, include=FALSE}
library(Hmisc)
library(datasets)
library(stats)
library(corrgram)
library(corrplot)
library(tidyverse)
```


# Data Set
- I'm going to stick with the HIX Compare Data we've been using so far
```{r loaddata}
load(file="plancounts.Rdata")
head(plancounts)
```

# Basic Statistics
```{r}
table(plancounts$YEAR,plancounts$METAL)
margin.table(table(plancounts$YEAR,plancounts$METAL),2)
testtbl <- prop.table(table(plancounts$YEAR,plancounts$METAL),1)
```

- resources
    - [quickR guide](https://www.statmethods.net/stats/index.html)
    - [R tutor guide](http://www.r-tutor.com/elementary-statistics)
    - [multivariate analysis guide](https://little-book-of-r-for-multivariate-analysis.readthedocs.io/en/latest/)
- correlations

```{r}
data("mtcars")
mtcars
cormat <- cor(mtcars)
rcorr(mtcars$cyl,mtcars$mpg, type="pearson")
rcorr(plancounts$avg27prem,plancounts$areaplans)
corrgram(mtcars, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.cor, text.panel=panel.txt)
```

- t-tests
```{r}
t.test(plancounts[plancounts$METAL==c("Gold","Silver"),]$avg27prem~plancounts[plancounts$METAL==c("Gold","Silver"),]$METAL)
```

- chi-square
```{r}
chisq.test(testtbl)
```

- wilcoxon rank sum
- mcnemar
- simple linear regression

# Multivariate Linear and Logistic Regressions
- next week we are going to do a review of regression assumptions
- then we will dive further into regression functions next week and the week after
- linear regression, output, using output from linear models
- logistic regression, attending to data types

```{r}
form <- formula(mpg~cyl+disp)

lm(form,data = mtcars)

logreg <- glm(am ~ disp + hp + cyl, data=mtcars, family=binomial(link="logit"))
summary(logreg)
```

