---
title: 'hsR Session 6: Regression Diagnostics & Assumptions in R'
author: "Nate Apathy"
date: "3/30/2018"
output: html_document
---

```{r pckgs&data, include=FALSE}
library(tidyverse)
library(Hmisc)
library(datasets)
library(stats)
library(corrgram)
library(corrplot)
library(tidyverse)
library(gvlma)

load(file="hixsamp.Rdata")
```

# Regression Assumptions

1. Model is linear in the parameters
2. Samples are independent and random (implies no serial correlation in the errors)
3. Zero conditional mean of the error (expected value of residual is 0)
4. No perfect collinearity in the regressors
5. Homoskedasticity of the error terms
6. Errors normally distributed
7. No serial correlation in the errors
8. Non-zero sample variance in x

# Resources
- [Quick-R Tutorial](https://www.statmethods.net/stats/rdiagnostics.html)
- [R Stats Detailed Walkthrough](http://r-statistics.co/Assumptions-of-Linear-Regression.html)

# Checking Specific Assumptions
*Some of these we don't need to "check," like linearity in the parameters. That has to do with model specification, and by doing a linear model we are constraining the parameters (ß values) to be linear. Similarly, the second assumption relates to sampling approach, and we will get to its implications in #7.*

*We will be doing these exercises on the hixsamp.Rdata we created earlier in the lessons.*

## First Step: Create a regression!
- Using the hixsamp data frame, let's make a basic model and then look at its adherence to the regression assumptions.
```{r}
# Build our regression
hixlm <- lm(PREMC2C30 ~ ST + as.factor(PLANTYPE) + METAL, data = hixsamp)

summary(hixlm)
```
- Here we've modeled the Premium amount for a family with 2 kids and 2 parents above age 30 as a function of state (California or Massachussets), plan type (PPO, HMO, etc), and Metal tier.

## Zero Conditional Mean
- This is satisfied if the average residual value is at (or so small to effectively be) zero.

```{r}
mean(hixlm$residuals)
# this appears to be satisfied because the value is so small
```














