---
title: "hsR Session 8: Gettin' Causal with Propensity Scores"
author: "Nate Apathy"
date: "4/20/2018"
output:
  html_document:
    highlight: tango
    theme: yeti
    df_print: paged
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r data&packages, include=FALSE}
library(tidyverse)
library(ggdag)
library(tableone)
library(Matching)
library(MatchIt)
```

# DAGs in R & Review of Confounding
- Confounders are variables that impact both your treatment and outcome
- Identifying and dealing with confounders is a major component of the work effort in developing a theoretical model, construct diagram, and model specification for analysis.
- Options in R
    - ggdag package - this is brand new and not well-developed, but it's what we have and can be useful!

## Example 1 - simple dag
```{r}
dagified <- dagify(x ~ z,
                   y ~ z,
                   exposure = "x",
                   outcome = "y")
tidy_dagitty(dagified)
ggdag(dagified, layout="circle")
```

## Example 2 - a more complex dag
```{r}
bigger_dag <- dagify(y ~ x + a + b,
                     x ~ a + b,
                     exposure = "x",
                     outcome = "y")

ggdag_paths(bigger_dag)

ggdag_parents(bigger_dag, "x")

ggdag_adjustment_set(bigger_dag)

```


# Propensity Scores & Matching

- examples that follow are from the [Crash Course in Causality on Coursera](https://www.coursera.org/learn/crash-course-in-causality/home/welcome) which I encourage everyone to do at some point
- it's a five week free course that goes into several components of causal inference with examples in R
- I've pulled these examples from that course
```{r}
# data from Crash Course in Causality (CCC) example
load(file="PS_ex.Rdata")
# object is psdata
head(psdata)
#covariates we will use (shorter list than you would use in practice)
xvars <- c("ARF","CHF","Cirr","colcan","Coma","lungcan","MOSF","sepsis",
         "age","female","meanbp1")

#look at a table 1, stratified on treatment status (0/1)
table1 <- CreateTableOne(vars=xvars, 
                         strata="treatment",
                         data=psdata, 
                         test=FALSE)
## include standardized mean difference (SMD)
print(table1, smd=TRUE)
# a few of these show imbalance (i.e. SMD > 0.1)
#tells you how much imbalance you might have
```

```{r}
##########################
#propensity score matching
#########################

#fit a propensity score model. logistic regression
# treatment is the outcome, with all of the covariates. family=binomial means the outcome is binary, which makes it carry out a logistic regression
psmodel <- glm(treatment ~ ARF +
                        CHF + 
                        Cirr + 
                        colcan + 
                        Coma + 
                        lungcan + 
                        MOSF + 
                        sepsis + 
                        age + 
                        female + 
                        meanbp1,
             family=binomial(),
             data=psdata)

#show coefficients etc
summary(psmodel)
#create propensity score
pscore <- psmodel$fitted.values
cbind(pscore,psmodel$model$treatment)
pscoreq <- ntile(pscore,4)

#newtable1
newpsdat <- data.frame(cbind(pscoreq,pscore,psmodel$model))

#fitted values are the values per subject of the fitted logistic model
#he graphs the propensity score, but doesn't include the code

#### greedy matching
# do greedy matching on logit(PS), using matching package
# caliper argument means 0.2 * the standard deviation of what you have X defined as
logit <- function(p) {log(p)-log(1-p)}
psmatch <- Match(Tr=psdata$treatment, 
                 M=1, X=logit(pscore), 
                 replace=FALSE, 
                 caliper=.2)
matched <- psdata[unlist(psmatch[c("index.treated","index.control")]), ]

#get standardized differences
matchedtab1<-CreateTableOne(vars=xvars, strata ="treatment", 
                            data=matched, test = FALSE)
print(matchedtab1, smd = TRUE)
#fewer subjects matched because of the restriction by the caliper
#previously, we tolerated wider distances
#we are now forcing better matches with the caliper
```

# Inverse Probability of Treatment Weighting (IPTW)


